|

### 2021-04-01 publication/manuscript.Rmd
iterate on supplement and manuscript

We include three control conditions to better understand the effectiveness of the investigated encoders:
The performance of a featureless learner (**FL** condition) was estimated as a conservative baseline for each dataset.
In regression problems, FL predicts the mean of the target variable in the training set for each observation in the test set.
In classification problems, the most frequent class of the target within the training set is predicted.
For each dataset, we also consider a random forest model without encoding (**none** condition), to compare the use of encoding methods in a random forest with a natural categorical splitting approach.
We use the random forest in ranger [@wright_2017] which provides efficient categorical feature support by ordering levels once before starting the tree growing algorithm [@wright_2019].
In the **remove** high cardinality features control condition, we remove features with a high number of levels above some threshold and use one-hot encoding (without collapsing rare levels) for the remaining features.
This condition reflects on whether including high cardinality features does indeed improve predictive performance.
Otherwise, the best encoding might just provide the least impairment compared to not including any high cardinality features.
We include an overview over available implementations in widely used machine learning frameworks for R and python in the supplementary material.

@@

Throughout our analysis, we use implementations of the various methods for the open-source statistical software R [@r_2018].
The pre-processing, as well as final ML algorithm described in \ref{sec:pipeline}, was trained and resampled using the `mlr` framework [@bischl_2016] together with the `batchtools` package [@lang_2017] to scale the benchmark analysis to HPC compute infrastructure.
All materials for this study, including reproducible code for this manuscript and the necessary result objects can be downloaded from our repository (see above).

@@

On a side note, we found that one-hot encoding usually gave a slightly better performance than dummy encoding.
This phenomenon has also been observed by @chiquet_2016.
@tutz_2016 commented that in generalized linear models "the naive combination of usual dummy coding with reference category and standard penalties is often a bad idea" (p.254) unless the reference category is of special interest.
In our benchmark, the reference category was chosen arbitrarily as is usually the case in ML.
Our results suggest that one-hot encoding is the better standard compared to dummy encoding when applying nonlinear regularized models like RF, GB or SVM with (high cardinal) categorical features.
We furthermore provide a more in-detail analysis of the performances of one-hot and dummy encoding in the supplementary material.

@@

Declarations of interest: none

>>>>

We include three control conditions to better understand the effectiveness of the investigated encoders:
The performance of a featureless learner (**FL** condition) was estimated as a conservative baseline for each dataset.
In regression problems, FL predicts the mean of the target variable in the training set for each observation in the test set.
In classification problems, the most frequent class of the target within the training set is predicted.
For each dataset, we also consider a random forest model without encoding (**none** condition), to compare the use of encoding methods in a random forest with a natural categorical splitting approach.
We use the random forest in ranger [@wright_2017] which provides efficient categorical feature support by ordering levels once before starting the tree growing algorithm [@wright_2019].
In the **remove** high cardinality features control condition, we remove features with a high number of levels above some threshold and use one-hot encoding (without collapsing rare levels) for the remaining features.
This condition reflects on whether including high cardinality features does indeed improve predictive performance.
Otherwise, the best encoding might just provide the least impairment compared to not including any high cardinality features.
We include an overview over available implementations in widely used machine learning frameworks for R and python in the Supplementary Material.

@@

Throughout our analysis, we use implementations of the various methods for the open-source statistical software R [@r_2018]. To enable a fair and reliable comparison in our study, we implemented all encoding methods on top of the `mlrCPO` package [@binder_2018].
The pre-processing, as well as final ML algorithm described in \ref{sec:pipeline}, was trained and resampled using the `mlr` framework [@bischl_2016] together with the `batchtools` package [@lang_2017] to scale the benchmark analysis to HPC compute infrastructure.
All materials for this study, including reproducible code for this manuscript and the necessary result objects can be downloaded from our **online repository** \footnote{\url{https://github.com/compstat-lmu/paper_2021_categorical_feature_encodings}}.

@@

On a side note, we found that one-hot encoding usually gave a slightly better performance than dummy encoding.
This phenomenon has also been observed by @chiquet_2016.
@tutz_2016 commented that in generalized linear models "the naive combination of usual dummy coding with reference category and standard penalties is often a bad idea" (p.254) unless the reference category is of special interest.
In our benchmark, the reference category was chosen arbitrarily as is usually the case in ML.
Our results suggest that one-hot encoding is the better standard compared to dummy encoding when applying nonlinear regularized models like RF, GB or SVM with (high cardinal) categorical features.
We furthermore provide a more in-detail analysis of the performances of one-hot and dummy encoding in the Supplementary Material.

@@

Declarations of interest {-}
==========
none

### 2021-04-01 publication/supplementary_material.Rmd
iterate on supplement and manuscript

---
title: |
  | Supplementary Material\:
  | Regularized target encoding outperforms traditional methods in supervised machine learning with high cardinality features
numbersections: true
csl: elsevier-harvard.csl
output: pdf_document
header-includes:
   - \usepackage{algorithm}
   - \usepackage{array}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{amsfonts}
   - \usepackage{mathtools}
   - \usepackage{bm}
   - \usepackage{tikz}
   - \usepackage{threeparttable}
   - \usepackage{booktabs}
   - \usetikzlibrary{trees,arrows, chains, fit, positioning, calc, shapes, shadows}
   - \setlength\parindent{0pt}
   - \usepackage{pdflscape}
   - \usepackage{float}
---

@@

Feature encoding techniques are implemented in several general preprocessing packages.
We provide an overview of existing implementations for R in Table \ref{tab:impls}.
For python, an extension to scikit-learn [@pedregosa_2011], *category encoders* [@will_mcginnis_2018_1157110] is available.
To enable a fair and reliable comparison in our study, we implemented all methods outlined above on top of the `mlrCPO` package.
All code can be found in our **online repository** \footnote{\url{https://github.com/compstat-lmu/paper_2021_categorical_feature_encodings}}.

\begin{table}
\centering
\begin{tabular}{|ccp{0.43\linewidth}|}
\hline
encoding method  & regularization & package \\
\hline
indicator        & --                &  \tt{stats} \newline \tt{embed}  \newline \tt{mlrCPO} \newline \tt{mlr3pipelines} \\
\hline
hash             & --                &  \tt{FeatureHashing} \newline \tt{embed} \\
\hline
impact           & smoothing         & \tt{embed} \newline \tt{mlrCPO} \newline \tt{mlr3pipelines} \\
\hline
impact           & cross-validation  & \tt{vtreat} \newline \texttt{mlr3pipelines} (via vtreat) \\
\hline
regularized impact (glmm) & --            & \tt{embed} \newline \tt{mlr3pipelines} \\
\hline
\end{tabular}
\caption{
  Overview over existing encoding implementations in R. Implementations can deviate from the algorithms described in section \ref{sec:methods} by minor implementation details.
  Indicator encoding encloses a variety of methods (one-hot, dummy, helmert encoding etc.)}
\label{tab:impls}
\end{table}

@@

@@

A frequently asked question is, whether one-hot encoding or dummy encoding is to be preferred. While dummy encoding clearly is to be preferred for non-regularized linear models in order to avoid multi-colinearity, the answer is less clear for general machine learning algorithms studied in our setting.

@@

```{r oneh-win, echo = FALSE, results="asis"}
oneh_win = indicator[Classes == 2 & !is.na(auc.test.mean_FALSE > auc.test.mean_TRUE),
  .(BinCl = sum(auc.test.mean_FALSE > auc.test.mean_TRUE)/.N*100), by = "lrn.id"][
indicator[Classes > 2 & !is.na(multiclass.aunu.test.mean_FALSE > multiclass.aunu.test.mean_TRUE),
  .(MultCl = sum(multiclass.aunu.test.mean_FALSE > multiclass.aunu.test.mean_TRUE)/.N*100), by = "lrn.id"],
    on = "lrn.id"][
indicator[Classes == 0 & !is.na(rmse.test.mean_FALSE < rmse.test.mean_TRUE),
  .(Regr = sum(rmse.test.mean_FALSE < rmse.test.mean_TRUE)/.N*100), by = "lrn.id"],
      on = "lrn.id"]
setnames(oneh_win, "lrn.id", "Learner")
setnames(oneh_win, "BinCl", "Binary Class")
setnames(oneh_win, "MultCl", "Multi Class")
setnames(oneh_win, "Regr", "Regression")
apa_table(oneh_win, digits = 0,
  caption = "Win Percentages of One-hot over Dummy Encoding",
  note = "Percentage of encoding conditions in which one-hot performs better than dummy per task setting.")
```

From the results shown in\ref{fig:oneh-win} we can observe, that for most algorithms and datasets, one-hot encoding is to be preferred over dummy-encoding.

>>>>

---
title: |
  | Supplementary Material\:
  | Regularized target encoding outperforms traditional methods in supervised machine learning with high cardinality features
numbersections: true
csl: elsevier-harvard.csl
bibliography: literature.bib
output: pdf_document
header-includes:
   - \usepackage{algorithm}
   - \usepackage{array}
   - \usepackage[noend]{algpseudocode}
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{amsfonts}
   - \usepackage{mathtools}
   - \usepackage{bm}
   - \usepackage{tikz}
   - \usepackage{threeparttable}
   - \usepackage{booktabs}
   - \usetikzlibrary{trees,arrows, chains, fit, positioning, calc, shapes, shadows}
   - \setlength\parindent{0pt}
   - \usepackage{pdflscape}
   - \usepackage{float}
---

@@

Feature encoding techniques are implemented in several general preprocessing packages.
We provide an overview of existing implementations for R in Table \ref{tab:impls}.
For python, an extension to scikit-learn [@pedregosa_2011], *category encoders* [@will_mcginnis_2018_1157110] is available.
To enable a fair and reliable comparison in our study, we implemented all methods outlined above on top of the `mlrCPO` package.
All code can be found in our **online repository** (<https://github.com/compstat-lmu/paper_2021_categorical_feature_encodings>).

\begin{table}[H]
\centering
\begin{tabular}{|ccp{0.43\linewidth}|}
\hline
encoding method  & regularization & package \\
\hline
indicator        & --                &  \tt{stats} \newline \tt{embed}  \newline \tt{mlrCPO} \newline \tt{mlr3pipelines} \\
\hline
hash             & --                &  \tt{FeatureHashing} \newline \tt{embed} \\
\hline
impact           & smoothing         & \tt{embed} \newline \tt{mlrCPO} \newline \tt{mlr3pipelines} \\
\hline
impact           & cross-validation  & \tt{vtreat} \newline \texttt{mlr3pipelines} (via vtreat) \\
\hline
regularized impact (glmm) & --            & \tt{embed} \newline \tt{mlr3pipelines} \\
\hline
\end{tabular}
\caption{
  Overview over existing encoding implementations in R. Implementations can deviate from the algorithms described in our manuscript by minor implementation details.
  Indicator encoding encloses a variety of methods (one-hot, dummy, helmert encoding etc.)}
\label{tab:impls}
\end{table}

@@

\newpage
\clearpage

@@

A frequently asked question is, whether one-hot encoding or dummy encoding is to be preferred. While dummy encoding clearly is commonly preferred for non-regularized linear models, the answer is less clear for general machine learning algorithms studied in our setting.

@@

```{r oneh-win, echo = FALSE, results="asis", fig.pos="!H"}
oneh_win = indicator[Classes == 2 & !is.na(auc.test.mean_FALSE > auc.test.mean_TRUE),
  .(BinCl = sum(auc.test.mean_FALSE > auc.test.mean_TRUE)/.N*100), by = "lrn.id"][
indicator[Classes > 2 & !is.na(multiclass.aunu.test.mean_FALSE > multiclass.aunu.test.mean_TRUE),
  .(MultCl = sum(multiclass.aunu.test.mean_FALSE > multiclass.aunu.test.mean_TRUE)/.N*100), by = "lrn.id"],
    on = "lrn.id"][
indicator[Classes == 0 & !is.na(rmse.test.mean_FALSE < rmse.test.mean_TRUE),
  .(Regr = sum(rmse.test.mean_FALSE < rmse.test.mean_TRUE)/.N*100), by = "lrn.id"],
      on = "lrn.id"]
setnames(oneh_win, "lrn.id", "Learner")
setnames(oneh_win, "BinCl", "Binary Class")
setnames(oneh_win, "MultCl", "Multi Class")
setnames(oneh_win, "Regr", "Regression")
apa_table(oneh_win, digits = 0,
  caption = "Win Percentages of One-hot over Dummy Encoding",
  note = "Percentage of encoding conditions in which one-hot performs better than dummy per task setting.")
```

From the results shown in Table 2 we can observe, that for most algorithms and datasets, one-hot encoding is to be preferred over dummy-encoding.
